---
title: "Understanding Near Earth Objects"
author: James Phillips & Ramtin Ranjpour
output-file: index.html
format:
  dashboard:
    theme: 
      - spacelab
execute:
  freeze: false
---

```{r}
#| label: load-packages
#| include: false

library(httr2) # For creating and submitting HTTP requests.
library(jsonlite) # For bidirectional mapping between JSON data and R data.
library(tidyverse) # For dplyr, purrr, tidyr, readr, and lubridate.
library(scales) # For aesthetic changes and methods for automatically determining breaks and labels for axes and legends.
library(knitr) # For report generation.
library(DT) # For the table.
library(htmltools) # For table interactivity.
library(flexdashboard) # For the table explainer.
library(bslib) # For the table explainer.
library(bsicons) # For the table explainer.

```

```{r}
#| label: get-parse-code
#| include: false

# DO NOT MODIFY WITHOUT CHECKING WITH JPL PARAMETERS.

sentry_get <- function(query = list()) {
  req <- request("https://ssd-api.jpl.nasa.gov/sentry.api") |> 
    req_url_query(!!!query) |> 
    req_user_agent("R (httr2); Sentry API client") |> 
    req_timeout(30)
  
  resp <- req |> req_perform()
  
  # Per JPL, throws an informative error on non-200.
  resp |> resp_check_status()
  
  txt <- resp |> resp_body_string()
  payload <- fromJSON(txt, simplifyVector = FALSE)
  
  # Handle API-level errors.
  if (!is.null(payload$error)) {
    stop(sprintf("Sentry API error: %s", payload$error), call. = FALSE)
  }
  
  # Version check recommended by JPL.
  ver <- payload$signature$version
  ver <- if (is.null(ver)) NA_character_ else as.character(ver)[1]
  
  if (!is.na(ver) && ver != "2.0") {
    warning(springt(
      "Sentry API signature version is %s (docs show 2.0). Format may differ.",
      ver
    ))
  }
  
  payload
}
```

```{r}
#| label: data-table-load
#| include: false

repair_fields <- function(fields, n_cols) {
  if (is.null(fields) || length(fields) != n_cols) {
    fields <- paste0("col_", seq_len(n_cols))
  } else {
    fields <- as.character(fields)
    bad <- is.na(fields) | trimws(fields) == ""
    fields[bad] <- paste0("col_", which(bad))
  }
  
  make.unique(fields)
}

as_sentry_table <- function(payload) {
  if (is.null(payload$data)) {
    stop("Payload has no $data element.")
  }
  
  dat <- payload$data
  
  if (is.list(dat) && length(dat) > 0 && is.list(dat[[1]])) {
    if (!is.null(names(dat[[1]])) && any(nzchar(names(dat[[1]])))) {
      out <- bind_rows(dat)
      return(as_tibble(out, .name_repair = "unique"))
    }
    
    if (!is.null(payload$fields)) {
      n_cols <- length(payload$fields)
      fields <- repair_fields(payload$fields, n_cols)
      
      rows <- map(dat, ~{
        x <- unlist(.x, use.names = FALSE)
        length(x) <- n_cols
        x
      })
      
      df <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
      names(df) <- fields
      return(as_tibble(df, .name_repair = "unique"))
    }
  }
  
  if (is.data.frame(dat)) {
    return(as_tibble(dat, .name_repair = "unique"))
  }
  
  out <- as_tibble(dat, .name_repair = "unique")
  return(out)
}

asteroid_summary <- sentry_get()
asteroid_table <- as_sentry_table(asteroid_summary)

asteroid_table_typed <- asteroid_table |> 
  mutate(
    ps_cum = parse_double(ps_cum),
    ps_max = parse_double(ps_max),
    ip = parse_double(ip),
    h = parse_double(h),
    ts_max = parse_double(ts_max),
    v_inf = parse_double(v_inf),
    diameter = parse_double(diameter),
    last_obs_jd = parse_double(last_obs_jd),
    n_imp = as.integer(n_imp),
    last_obs = as.Date(last_obs)
  ) |> 
  mutate(designation = des)

asteroid_table_typed$last_obs_jd <- NULL
asteroid_table_typed$id <- NULL
asteroid_table_typed$h <- NULL
asteroid_table_typed$des <- NULL

reorder <- c(
  "fullname", "last_obs", "diameter", "ip", "n_imp", "v_inf", "range", "ps_cum", "ps_max", "ts_max"
)

asteroid_table_typed <- asteroid_table_typed |> 
  select(any_of(reorder), everything())

# Use `asteroid_table_typed` dataframe for all visualizations. Modify as needed, but do NOT use `asteroid_table` or `asteroid_summary`.
```

```{r}
#| label: sentry-summary-s
#| include: false

s_payload <- sentry_get(list())

sentry_summary_tbl <- bind_rows(s_payload$data) |> 
  as_tibble(.name_repair = "unique") |> 
  mutate(across(everything(), ~na_if(as.character(.x), "")))

objects_after_2126_tbl <- sentry_summary_tbl |> 
  mutate(
    end_year = suppressWarnings(as.integer(str_extract(range, "(?<=-)\\d{4}$")))
  ) |> 
  filter(!is.na(end_year), end_year > 2126) |> 
  arrange(end_year, fullname) |> 
  select(fullname, range)
```

# About This Project

At any given moment, experts in space science are tracking over 2,000 near-Earth objects (NEOs) that could pose a potential risk to life on Earth due to their orbital trajectories, estimated diameters, and projected velocities. [NASA's JPL Sentry System](https://cneos.jpl.nasa.gov/sentry/) is an Earth impact monitoring system designed for comprehensive observation and collection of data regarding these objects. But how do people not intimately familiar with space science concepts engage with this information? This project seeks to make data regarding possible impact events more approachable to the everyday person, presenting different elements of Sentry's data in ways that better contextualize the actual risk of currently monitored NEOs, the range in which potential impacts might occur, and the scale of risk currently associated with said objects.

# Potential Impacts Timeline

::: sidebar
## About This Graph

This graph shows a timeline of possible asteroid impacts based on current data from Sentry.

Each bubble represents a 10-year time window, beginning with the current decade (2020-2029). Bubble sizes increase based on overall number of possible asteroids with impact risks in those timeframes.

**A note**: No data is provided after the year 2126 as Sentry only provides tracking information for objects with possible impacts in a 100-year window. Anomalies that do appear are excluded from this visualization but remain in the original dataset potentially as the result of prominent media presence, as is the case with the asteroid 101955 Bennu. The objects are as follows:

```{r}
#| label: after-2126-table
#| echo: false
#| results: asis

if (nrow(objects_after_2126_tbl) == 0) {
  cat("<div style='font-size:0.9em;opacity:0.8;'>None.</div>")
} else {
  knitr::kable(
    objects_after_2126_tbl,
    format = "html",
    col.names = c("Object", "Impact range"),
    escape = TRUE
  )
}
```

`r nrow(objects_after_2126_tbl)`

## Data Recency

```{r, echo=FALSE, results='asis'}
render_utc <- format(Sys.time(), tz = "UTC", "%Y-%m-%dT%H:%M:%SZ")

cat(paste0(
  "<span id='last-updated' data-utc='", render_utc, "'></span><br>",
  "<span id='next-refresh'></span>"
))
```

```{=html}
<script>
  (function () {
    // Read the UTC timestamp embedded by R
    const el = document.getElementById('last-updated');
    if (!el) return;
    
    const utc = el.getAttribute('data-utc');
    const last = new Date(utc);
    
    // Compute next refresh (daily at 14:00 UTC)
    const nextUtc = new Date();
    nextUtc.setUTCHours(14, 0, 0, 0);
    if (new Date() >= nextUtc) {
      nextUtc.setUTCDate(nextUtc.getUTCDate() + 1);
    }
    
    // Format in the viewer's local timezone
    const fmt = new Intl.DateTimeFormat(undefined, {
    dateStyle: 'long',
    timeStyle: 'short'
    });
    
    el.textContent = "This data was last updated on " + fmt.format(last) + ".";
    document.getElementById('next-refresh').textContent = 
      "Next scheduled update: " + fmt.format(nextUtc) + ".";
  })();
</script>
```
:::

## {.tabset}

### Updated

```{r}
#| label: impacts-timeline-bubble-plot-update
#| warning: false

bin_width <- 10

impacts_parsed <- asteroid_table_typed |> 
  mutate(
    range = str_trim(range),
    range_ok = str_detect(range, "^\\d{4}\\s*-\\s*\\d{4}$")
  ) |> 
  filter(range_ok) |> 
  separate(
    range,
    into = c("start_year", "end_year"),
    sep = "\\s*-\\s*",
    convert = TRUE,
    remove = FALSE
  ) |> 
  filter(!is.na(start_year), !is.na(end_year), end_year >= start_year)

bin_start <- floor(min(impacts_parsed$start_year) / bin_width) * bin_width
bin_end <- ceiling(max(impacts_parsed$end_year) / bin_width) * bin_width
  
impacts_timeline <- tibble(
  bin_lo = seq(bin_start, bin_end - bin_width, by = bin_width),
  bin_hi = bin_lo + bin_width - 1
) |> 
  rowwise() |> 
  mutate(
    n_objects = sum(impacts_parsed$start_year <= bin_hi & impacts_parsed$end_year >= bin_lo),
    bin_mid = (bin_lo + bin_hi) / 2,
    bin_label = paste0(bin_lo, "-", bin_hi)
  ) |> 
  ungroup() |> 
  filter(n_objects > 0) |> 
  filter(bin_hi <= 2129) |> 
  arrange(bin_mid) |> 
  mutate(
    y = 0,
    side = if_else(row_number() %% 2 == 0, "below", "above"),
    y_label = if_else(side == "above", 0.30, -0.30),
    y_line_end = if_else(side == "above", 0.27, -0.27),
    label = paste0(bin_label, "\n(n = ", n_objects, ")")
  )

xmin <- min(impacts_timeline$bin_lo)
xmax <- max(impacts_timeline$bin_hi)

ggplot(impacts_timeline, aes(x = bin_mid, y = y)) +
  geom_segment(
    aes(x = xmin, xend = xmax, y = 0, yend = 0),
    inherit.aes = FALSE,
    linewidth = 1.1,
    arrow = arrow(length = unit(0.22, "cm"), type = "closed")
  ) +
  geom_segment(
    aes(xend = bin_mid, yend = y_line_end),
    linewidth = 0.35, alpha = 0.35
  ) +
  geom_point(
    aes(size = n_objects),
    shape = 21,
    fill = "#56B4E9",
    color = "#56B4E9",
    alpha = 0.65
  ) +
  geom_text(
    aes(y = y_label, label = label),
    size = 2.5,
    lineheight = 0.95,
    angle = 45,
    hjust = 0.5,
    vjust = 0.5) +
  scale_x_continuous(
    name = NULL,
    breaks = NULL,
    labels = NULL,
    expand = expansion(mult = c(0.05, 0.10))
  ) +
  scale_y_continuous(
    name = NULL,
    breaks = NULL
  ) +
  scale_size(range = c(5, 30), guide = "none") +
  coord_cartesian(clip = "off") +
  theme_void(base_size = 14) +
  theme(
    plot.margin = margin(25, 25, 25, 25),
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20))
    ) +
  labs(
    title = "Timeline of Possible NEO Impacts Between 2026 and 2129"
  )
```


### Original

```{r}
#| label: impacts-timeline-bubble-plot-orig
#| warning: false

bin_width <- 10

impacts_parsed <- asteroid_table_typed |> 
  mutate(
    range = str_trim(range),
    range_ok = str_detect(range, "^\\d{4}\\s*-\\s*\\d{4}$")
  ) |> 
  filter(range_ok) |> 
  separate(
    range,
    into = c("start_year", "end_year"),
    sep = "\\s*-\\s*",
    convert = TRUE,
    remove = FALSE
  ) |> 
  filter(!is.na(start_year), !is.na(end_year), end_year >= start_year)

bin_start <- floor(min(impacts_parsed$start_year) / bin_width) * bin_width
bin_end <- ceiling(max(impacts_parsed$end_year) / bin_width) * bin_width
  
impacts_timeline <- tibble(
  bin_lo = seq(bin_start, bin_end - bin_width, by = bin_width),
  bin_hi = bin_lo + bin_width - 1
) |> 
  rowwise() |> 
  mutate(
    n_objects = sum(impacts_parsed$start_year <= bin_hi & impacts_parsed$end_year >= bin_lo),
    bin_mid = (bin_lo + bin_hi) / 2,
    bin_label = paste0(bin_lo, "-", bin_hi)
  ) |> 
  ungroup() |> 
  filter(n_objects > 0) |> 
  arrange(bin_mid) |> 
  mutate(
    y = 0,
    side = if_else(row_number() %% 2 == 0, "below", "above"),
    y_label = if_else(side == "above", 0.30, -0.30),
    y_line_end = if_else(side == "above", 0.27, -0.27),
    label = paste0(bin_label, "\n(n = ", n_objects, ")")
  )

xmin <- min(impacts_timeline$bin_lo)
xmax <- max(impacts_timeline$bin_hi)

ggplot(impacts_timeline, aes(x = bin_mid, y = y)) +
  geom_segment(
    aes(x = xmin, xend = xmax, y = 0, yend = 0),
    inherit.aes = FALSE,
    linewidth = 1.1,
    arrow = arrow(length = unit(0.22, "cm"), type = "closed")
  ) +
  geom_segment(
    aes(xend = bin_mid, yend = y_line_end),
    linewidth = 0.35, alpha = 0.35
  ) +
  geom_point(
    aes(size = n_objects),
    shape = 21,
    fill = "#56B4E9",
    color = "#56B4E9",
    alpha = 0.65
  ) +
  geom_text(
    aes(y = y_label, label = label),
    size = 2.5,
    lineheight = 0.95,
    angle = 45,
    hjust = 0.5,
    vjust = 0.5) +
  scale_x_continuous(
    name = NULL,
    breaks = NULL,
    labels = NULL,
    expand = expansion(mult = c(0.05, 0.10))
  ) +
  scale_y_continuous(
    name = NULL,
    breaks = NULL
  ) +
  scale_size(range = c(4, 20), guide = "none") +
  coord_cartesian(clip = "off") +
  theme_void(base_size = 14) +
  theme(
    plot.margin = margin(25, 25, 25, 25),
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20))
    ) +
  labs(
    title = "Timeline of Possible NEO Impacts Between 2026 and 2289"
  )
```

# Palermo Scale Risk Distribution

```{r}
#| label: palermo-bar-chart
#| warning: false
#| fig-cap: "Distribution of monitored asteroids across Palermo Scale risk categories"

asteroid_risk <- asteroid_table_typed |>
  mutate(
    ps_cum_cat = case_when(
      ps_cum > 0 ~ ">0",
      ps_cum <= 0 & ps_cum > -2 ~ ">0 to =-2",
      ps_cum <= -2 & ps_cum > -4 ~ ">-2 to =-4",
      ps_cum <= -4 & ps_cum > -6 ~ ">-4 to =-6",
      ps_cum <= -6 ~ "<=-6",
      TRUE ~ NA_character_
    ),
    ps_max_cat = case_when(
      ps_max > 0 ~ ">0",
      ps_max <= 0 & ps_max > -2 ~ ">0 to =-2",
      ps_max <= -2 & ps_max > -4 ~ ">-2 to =-4",
      ps_max <= -4 & ps_max > -6 ~ ">-4 to =-6",
      ps_max <= -6 ~ "<=-6",
      TRUE ~ NA_character_
    )
  )

risk_long <- asteroid_risk |>
  select(ps_cum_cat, ps_max_cat) |>
  pivot_longer(
    cols = everything(),
    names_to = "scale",
    values_to = "category"
  ) |>
  filter(!is.na(category))

risk_long$category <- factor(
  risk_long$category,
  levels = c(">0", ">0 to =-2", ">-2 to =-4", ">-4 to =-6", "<=-6")
)

risk_counts <- risk_long |>
  count(scale, category)

ggplot(risk_counts,
       aes(x = category, y = n, fill = scale)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = comma) +
  labs(
    x = "Palermo Scale Risk Category",
    y = "Number of Asteroids",
    fill = "Palermo Scale",
    title = "Relative Risk Distribution of Monitored Near-Earth Asteroids",
    subtitle = "Comparison of cumulative (ps_cum) and maximum (ps_max) Palermo Scale values"
  ) +
  theme_minimal(base_size = 12)

```

# Sentry Objects Table

```{r}
#| label: sentry-object-table
#| echo: false

obj_tbl <- asteroid_table_typed |> 
  transmute(
    designation,
    fullname,
    range,
    n_imp,
    ip,
    v_inf,
    diameter,
    ps_cum,
    ps_max,
    ts_max,
    last_obs
  )

datatable(
  obj_tbl,
  rownames = FALSE,
  selection = "single",
  colnames = c(
    "Designation",
    "Object",
    "Impact Year Range",
    "Potential Impacts",
    "Impact Probability (cumulative)",
    "Velocity (km/s)",
    "Est. Diameter (km)",
    "Palermo Scale (cum.)",
    "Palermo Scale (max.)",
    "Torino Scale (max.)",
    "Last Observed"
  ),
  options = list(
    pageLength = 10,
    paging = TRUE,
    info = TRUE,
    searching = TRUE,
    lengthMenu = list(c(10,25,50,100), c("10", "25", "50", "100")),
    lengthChange = TRUE,
    autoWidth = TRUE,
    scrollX = TRUE,
    dom = "<'dt-top'lfr>t<'dt-bottom'ip>",
    pagingType = "simple_numbers",
    columnDefs = list(list(targets = 0, visible = FALSE))
  ),
  callback = DT::JS("
    table.on('click.dt', 'tr', function() {
      var row = table.row(this);
      if (!row || !row.data()) return;
      
      var data = row.data();
      var des = data[0];
      
      var panel = document.getElementById('sentry-details');
      panel.innerHTML = '<div style=\"opacity:.8;\">Loading details for <b>' + des + '</b>...</div>';
      
      var url = 'https://ssd-api.jpl.nasa.gov/sentry.api?des=' + encodeURIComponent(des);
      
      fetch(url)
        .then(r => r.json())
        .then(payload => {
          if (payload.error) {
            panel.innerHTML = '<div style=\"color:#b00020;\">API error: ' + payload.error + '</div>';
            return;
          }
          
          var s = (payload.summary && payload.summary.length) ? payload.summary[0] : null;
          if (!s) {
            panel.innerHTML = '<div>No summary available for ' + des + '.</div>';
            return;
          }
          
          var html = '';
          html += '<h3 style=\"margin-top:0;\">' + (s.fullname || des) + '</h3>';
          html += '<div style=\"display:grid;grid-template-columns:1fr 1fr;gap:8px;\">';
          html += '<div><b>Year range:</b> ' + (s.range ?? '') + '</div>';
          html += '<div><b>Potential impacts:</b> ' + (s.n_imp ?? '') + '</div>';
          html += '<div><b>Cumulative IP:</b> ' + (s.ip ?? '') + '</div>';
          html += '<div><b>V∞ (km/s):</b> ' + (s.v_inf ?? '') + '</div>';
          html += '<div><b>Diameter (km):</b> ' + (s.diameter ?? '') + '</div>';
          html += '<div><b>Torino max:</b> ' + (s.ts_max ?? '') + '</div>';
          html += '<div><b>Palermo cum:</b> ' + (s.ps_cum ?? '') + '</div>';
          html += '<div><b>Palermo max:</b> ' + (s.ps_max ?? '') + '</div>';
          html += '</div>';
          
          var impacts = payload.data || [];
          html += '<h4 style=\"margin-top:14px;\">Impact events (sample)</h4>';
          
          if (!impacts.length) {
            html += '<div style=\"opacity:.8;\">No impact-event rows returned.</div>';
          } else {
            var head = Object.keys(impacts[0]);
            var showCols = ['date', 'ip', 'ps', 'ts', 'sigma', 'energy']; // best-effort
            var cols = head.filter(h => showCols.includes(h)).slice(0, 6);
            if (!cols.length) cols = head.slice(0, 6);

            html += '<table style=\"width:100%;border-collapse:collapse;\">';
            html += '<thead><tr>' + cols.map(c => '<th style=\"text-align:left;border-bottom:1px solid #ddd;padding:4px;\">' + c + '</th>').join('') + '</tr></thead>';
            html += '<tbody>';

            impacts.slice(0, 10).forEach(row => {
              html += '<tr>' + cols.map(c => '<td style=\"border-bottom:1px solid #f0f0f0;padding:4px;\">' + (row[c] ?? '') + '</td>').join('') + '</tr>';
            });

            html += '</tbody></table>';
            html += '<div style=\"margin-top:6px;opacity:.75;font-size:.9em;\">Showing 10 rows; API returned ' + impacts.length + '.</div>';
          }
          
          panel.innerHTML = html;
        })
        .catch(err => {
          panel.innerHTML = '<div style=\"color:#b00020;\">Fetch failed: ' + err + '</div>';
        });
      });
  ")
)
```

# Understanding the Table {scrolling="true"}

::: {.card title="Case Study: 2024 YR4" fill="false"}

Each row in NASA's sentry table represents a *near-Earth Object (NEO)* currently monitored by Sentry that presents any statistical chance of impacting the Earth. To illustrate what each column and data point represents, we will use the asteroid **2024 YR4** as an example, which briefly had a **3.1%** chance of hitting Earth before further calculations reduced that risk to **0.00081%**, generating [significant media attention](https://www.esa.int/Space_Safety/Planetary_Defence/ESA_actively_monitoring_near-Earth_asteroid_2024_YR4) as a result.

:::

::: {.card title="Designation Breakdown"}

:::

##

```{r}
#| label: year-detect

value_box(
  title = "Year Detected",
  value = "2024",
  showcase = bs_icon("calendar"),
  showcase_layout = "left center",
  theme = value_box_theme(bg = "#20438e"),
  height = "150px"
)
```

```{r}
#| label: half-month

value_box(
  title = "Half-Month Code",
  value = "Y",
  showcase = bs_icon("type"),
  theme = value_box_theme(bg = "#e6f2fd"),
  height = "150px",
  p(
    "2nd Half of December",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: sequence-sub

value_box(
  title = "Sequence Subscript",
  value = "R4",
  showcase = bs_icon("list-ol"),
  theme = value_box_theme(bg = "#20438e"),
  height = "150px",
  p(
    "117th Designation That Half-Month",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: peak-prob

value_box(
  title = "Peak Probability",
  value = "3.1%",
  showcase = bs_icon("exclamation-triangle"),
  theme = value_box_theme(bg = "#e6f2fd"),
  height = "150px",
  p(
    "Highest Recorded (Jan 2025)",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: current-prob

value_box(
  title = "Current Probability",
  value = ".00081%",
  showcase = bs_icon("percent"),
  theme = value_box_theme(bg = "#20438e"),
  height = "150px",
  p(
    "Revised Probability (Mar 2026)",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

##

::: {.card title="Identification Data"}

:::

##

```{r}
#| label: object-des

value_box(
  title = NULL,
  value = div(
    "Object",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("tag"),
  theme = value_box_theme(bg = "#20438e"),
  height = "150px",
  p(
    "Provisional designation: Year + half-month code + sequence subscript",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: last-obs

value_box(
  title = NULL,
  value = div(
    "Last Observed",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("eye"),
  theme = value_box_theme(bg = "#e6f2fd"),
  height = "150px",
  p(
    "Most recent observation date",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: est-diam

value_box(
  title = NULL,
  value = div(
    "Estimated Diameter",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("arrows-angle-expand"),
  theme = value_box_theme(bg = "#20438e"),
  height = "150px",
  p(
    "Best guess based on shape and light reflection intensity",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

##

::: {.card title="Orbit & Impact Data"}

:::

##

```{r}
#| label: year-range

value_box(
  title = NULL,
  value = div(
    "Impact Year Range",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("calendar"),
  theme = value_box_theme(bg = "#20438e"),
  height = "150px",
  p(
    "Years in which impact could occur",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: pot-impacts

value_box(
  title = NULL,
  value = div(
    "Potential Impacts",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("circle"),
  theme = value_box_theme(bg = "#e6f2fd"),
  height = "150px",
  p(
    "Possible impacts throughout year range",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: impact-prob

value_box(
  title = NULL,
  value = div(
    "Impact Probability",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("dice-1"),
  theme = value_box_theme(bg = "#20438e"),
  height = "150px",
  p(
    "Cumulative chance across all passes",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

```{r}
#| label: velocity-inf

value_box(
  title = NULL,
  value = div(
    "Velocity (V∞)",
    style = "font-size: 2rem; font-weight: 600;"
  ),
  showcase = bs_icon("speedometer2"),
  theme = value_box_theme(bg = "#e6f2fd"),
  height = "150px",
  p(
    "Speed without major gravitational pull",
    style = "font-size: 0.75rem; font-style: italic; margin-bottom: 0;"
  )
)
```

##

::: {.card title="Reference Scales"}

:::

##

::: {.card title="Palermo Scale"}

```{r}
#| label: palermo-scale

palermo <- data.frame(
  value   = c("< -2", "-2 to -1", "-1 to 0", "0", "> 0"),
  meaning = c("Less than 1% of background risk", "1-10% of background risk",
              "10-100% of background risk",      "Equal to background risk",
              "Exceeds all background risk"),
  level   = c("Minimal concern", "Worth monitoring", "Careful attention",
              "Serious attention", "High priority"),
  bg      = c("#2d6a4f", "#52b788", "#d4a017", "#e07b20", "#c0392b"),
  fg      = c("#ffffff", "#ffffff", "#ffffff", "#ffffff", "#ffffff")
)

grad_bar <- div(
  style = "height:14px; border-radius:6px; margin-bottom:10px;
           background: linear-gradient(90deg,
             #2d6a4f 0%, #52b788 25%, #d4a017 50%, #e07b20 75%, #c0392b 100%);"
)

tick_labels <- tags$table(
  style = "width:100%; border-collapse:collapse; margin-bottom:12px;",
  tags$tr(
    tags$td(style = "width:20%; text-align:left; font-size:11px; color:#666;", "< -2"),
    tags$td(style = "width:20%; text-align:center; font-size:11px; color:#666;", "-2"),
    tags$td(style = "width:20%; text-align:center; font-size:11px; color:#666;", "-1"),
    tags$td(style = "width:20%; text-align:center; font-size:11px; color:#666;", "0"),
    tags$td(style = "width:20%; text-align:right; font-size:11px; color:#666;", "> 0")
  )
)

rows <- lapply(seq_len(nrow(palermo)), function(i) {
  tags$tr(
    tags$td(style = paste0("background:", palermo$bg[i], "; color:", palermo$fg[i],
                           "; font-weight:600; padding:6px 10px; border-radius:3px;
                           text-align:center; white-space:nowrap;"),
            palermo$value[i]),
    tags$td(style = "padding:6px 10px; font-size:13px;", palermo$meaning[i]),
    tags$td(style = paste0("padding:6px 10px; font-size:13px; font-weight:600; color:",
                           palermo$bg[i], ";"), palermo$level[i])
  )
})

tagList(
  tags$p(style = "font-size:11px; color:#888; margin-bottom:10px; font-style:italic;",
         "Logarithmic · Technical Audience"),
  grad_bar,
  tick_labels,
  tags$table(style = "width:100%; border-collapse:collapse;", tags$tbody(tagList(rows))),
  tags$p(style = "font-size:12px; color:#888; margin-top:10px; font-style:italic;",
         HTML("Reported as <b>Cumulative</b> (all passes) and <b>Maximum</b> (worst single pass).
               Nothing currently exceeds -2."))
)
```

:::


::: {.card title="Torino Scale"}

```{r}
#| label: torino-scale

torino <- data.frame(
  score   = c("0", "1", "2-4", "5-7", "8-10"),
  label   = c("No Hazard", "Normal", "Attention", "Threatening", "Certain"),
  desc    = c("Burns up in atmosphere or no risk",
              "Routine pass, no discernible danger",
              "Warrants monitoring by astronomers",
              "Potentially unprecedented global threat",
              "Collision certain; size determines severity"),
  bg      = c("#adb5bd", "#2d6a4f", "#d4a017", "#e07b20", "#c0392b"),
  fg      = c("#ffffff", "#ffffff", "#ffffff", "#ffffff", "#ffffff")
)

torino_cols <- c("#adb5bd","#2d6a4f","#52b788","#7bc67e",
                 "#d4a017","#d4a017","#e07b20","#e07b20",
                 "#c0392b","#c0392b","#8b0000")

cell_strip <- tags$table(
  style = "width:100%; border-collapse:collapse; margin-bottom:12px; table-layout:fixed;",
  tags$tr(
    lapply(0:10, function(i) {
      tags$td(
        style = paste0(
          "background:", torino_cols[i + 1], "; color:#fff; font-weight:700; ",
          "font-size:12px; text-align:center; padding:7px 0; ",
          if (i == 0) "border-top-left-radius:5px; border-bottom-left-radius:5px;" else "",
          if (i == 10) "border-top-right-radius:5px; border-bottom-right-radius:5px;" else ""
        ),
        as.character(i)
      )
    })
  )
)

rows <- lapply(seq_len(nrow(torino)), function(i) {
  tags$tr(
    tags$td(style = paste0("background:", torino$bg[i], "; color:", torino$fg[i],
                           "; font-weight:700; padding:6px 10px; border-radius:3px;
                           text-align:center; white-space:nowrap;"),
            torino$score[i]),
    tags$td(style = paste0("padding:6px 10px; font-size:13px; font-weight:600; color:",
                           torino$bg[i], ";"), torino$label[i]),
    tags$td(style = "padding:6px 10px; font-size:13px;", torino$desc[i])
  )
})

tagList(
  tags$p(style = "font-size:11px; color:#888; margin-bottom:10px; font-style:italic;",
         "Integer 0-10 · Public Communication · Impacts < 100 Years Away"),
  cell_strip,
  tags$table(style = "width:100%; border-collapse:collapse;", tags$tbody(tagList(rows))),
  tags$p(style = "font-size:12px; color:#888; margin-top:10px; font-style:italic;",
         HTML("No known object currently exceeds a Torino score of <b>1</b>."))
)
```

:::