---
title: "Understanding Near Earth Objects"
author: James Phillips & Ramtin Ranjpour
output-file: index.html
format:
  dashboard:
    theme: 
      - spacelab
execute:
  freeze: false
---

```{r}
#| label: load-packages
#| include: false

library(httr2) # For creating and submitting HTTP requests.
library(jsonlite) # For bidirectional mapping between JSON data and R data.
library(tidyverse) # For dplyr, purrr, tidyr, readr, and lubridate.
library(scales) # For aesthetic changes and methods for automatically determining breaks and labels for axes and legends.
library(knitr) # For report generation.
library(DT) # For the table.
library(htmltools) # For table interactivity.

```

```{r}
#| label: get-parse-code
#| include: false

# DO NOT MODIFY WITHOUT CHECKING WITH JPL PARAMETERS.

sentry_get <- function(query = list()) {
  req <- request("https://ssd-api.jpl.nasa.gov/sentry.api") |> 
    req_url_query(!!!query) |> 
    req_user_agent("R (httr2); Sentry API client") |> 
    req_timeout(30)
  
  resp <- req |> req_perform()
  
  # Per JPL, throws an informative error on non-200.
  resp |> resp_check_status()
  
  txt <- resp |> resp_body_string()
  payload <- fromJSON(txt, simplifyVector = FALSE)
  
  # Handle API-level errors.
  if (!is.null(payload$error)) {
    stop(sprintf("Sentry API error: %s", payload$error), call. = FALSE)
  }
  
  # Version check recommended by JPL.
  ver <- payload$signature$version
  ver <- if (is.null(ver)) NA_character_ else as.character(ver)[1]
  
  if (!is.na(ver) && ver != "2.0") {
    warning(springt(
      "Sentry API signature version is %s (docs show 2.0). Format may differ.",
      ver
    ))
  }
  
  payload
}
```

```{r}
#| label: data-table-load
#| include: false

repair_fields <- function(fields, n_cols) {
  if (is.null(fields) || length(fields) != n_cols) {
    fields <- paste0("col_", seq_len(n_cols))
  } else {
    fields <- as.character(fields)
    bad <- is.na(fields) | trimws(fields) == ""
    fields[bad] <- paste0("col_", which(bad))
  }
  
  make.unique(fields)
}

as_sentry_table <- function(payload) {
  if (is.null(payload$data)) {
    stop("Payload has no $data element.")
  }
  
  dat <- payload$data
  
  if (is.list(dat) && length(dat) > 0 && is.list(dat[[1]])) {
    if (!is.null(names(dat[[1]])) && any(nzchar(names(dat[[1]])))) {
      out <- bind_rows(dat)
      return(as_tibble(out, .name_repair = "unique"))
    }
    
    if (!is.null(payload$fields)) {
      n_cols <- length(payload$fields)
      fields <- repair_fields(payload$fields, n_cols)
      
      rows <- map(dat, ~{
        x <- unlist(.x, use.names = FALSE)
        length(x) <- n_cols
        x
      })
      
      df <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
      names(df) <- fields
      return(as_tibble(df, .name_repair = "unique"))
    }
  }
  
  if (is.data.frame(dat)) {
    return(as_tibble(dat, .name_repair = "unique"))
  }
  
  out <- as_tibble(dat, .name_repair = "unique")
  return(out)
}

asteroid_summary <- sentry_get()
asteroid_table <- as_sentry_table(asteroid_summary)

asteroid_table_typed <- asteroid_table |> 
  mutate(
    ps_cum = parse_double(ps_cum),
    ps_max = parse_double(ps_max),
    ip = parse_double(ip),
    h = parse_double(h),
    ts_max = parse_double(ts_max),
    v_inf = parse_double(v_inf),
    diameter = parse_double(diameter),
    last_obs_jd = parse_double(last_obs_jd),
    n_imp = as.integer(n_imp),
    last_obs = as.Date(last_obs)
  ) |> 
  mutate(designation = des)

asteroid_table_typed$last_obs_jd <- NULL
asteroid_table_typed$id <- NULL
asteroid_table_typed$h <- NULL
asteroid_table_typed$des <- NULL

reorder <- c(
  "fullname", "last_obs", "diameter", "ip", "n_imp", "v_inf", "range", "ps_cum", "ps_max", "ts_max"
)

asteroid_table_typed <- asteroid_table_typed |> 
  select(any_of(reorder), everything())

# Use `asteroid_table_typed` dataframe for all visualizations. Modify as needed, but do NOT use `asteroid_table` or `asteroid_summary`.
```

```{r}
#| label: sentry-summary-s
#| include: false

s_payload <- sentry_get(list())

sentry_summary_tbl <- bind_rows(s_payload$data) |> 
  as_tibble(.name_repair = "unique") |> 
  mutate(across(everything(), ~na_if(as.character(.x), "")))

objects_after_2126_tbl <- sentry_summary_tbl |> 
  mutate(
    end_year = suppressWarnings(as.integer(str_extract(range, "(?<=-)\\d{4}$")))
  ) |> 
  filter(!is.na(end_year), end_year > 2126) |> 
  arrange(end_year, fullname) |> 
  select(fullname, range)
```

# About This Project

At any given moment, experts in space science are tracking over 2,000 near-Earth objects (NEOs) that could pose a potential risk to life on Earth due to their orbital trajectories, estimated diameters, and projected velocities. [NASA's JPL Sentry System](https://cneos.jpl.nasa.gov/sentry/) is an Earth impact monitoring system designed for comprehensive observation and collection of data regarding these objects. But how do people not intimately familiar with space science concepts engage with this information? This project seeks to make data regarding possible impact events more approachable to the everyday person, presenting different elements of Sentry's data in ways that better contextualize the actual risk of currently monitored NEOs, the range in which potential impacts might occur, and the scale of risk currently associated with said objects.

## Some Data Explainers

While some of the data in this project may be self-explanatory, we want to be exhaustive in providing explanations for any possible data used so as to reduce confusion and create better accessibility with this data. Each item you might encounter while exploring these visualizations, or the data table itself, is listed below:

[Object]{style="font-size: 18px;font-weight: bold;"}: The assigned provisional designation, or informal name, of the NEO. Typically, NEOs are assigned a designation based on when the object was initially monitored. This follows a standard formula: the *year of discovery* followed by a two-letter code indicating the *half-month of discovery*, including a subscript to help identify *what position that object is in the given half-month in terms of order of discovery*. Our example: asteroid 2024 YR4, which gained [significant media traction](https://www.esa.int/Space_Safety/Planetary_Defence/ESA_actively_monitoring_near-Earth_asteroid_2024_YR4) in early 2025 due to an anomalously high chance (3.1%) of hitting Earth in 2032. It's designation was determined because it was detected in 2024 (2024), in the second half of the month of December (Y), and was the 117th provisional designation assigned in that half-month period (R4). More information on provisional designations can be found [here](https://en.wikipedia.org/wiki/Provisional_designation_in_astronomy).

[Impact Year Range]{style="font-size: 18px;font-weight: bold;"}: The range of years in which an impact could occur. These ranges tend to fluctuate as further observations of the object are made.

[Potential Impacts]{style="font-size: 18px;font-weight: bold;"}: The number of possible impacts the object could have in the given year range. This is based on how many "passes" the object will make by Earth in its orbit (as in, each orbit could present a possible impact risk, but not always).

[Impact Probability (cumulative)]{style="font-size: 18px;font-weight: bold;"}: Probability that the possible impact will occur. Calculating this probability is complex and depends on a number of assumptions that are difficult to verify. For these reasons, the stated probability can easily be inaccurate by a factor of a few, and occasionally by a factor of ten or more. Returning to our example of 2024 YR4, at the time of this writing (March 3, 2026), 2024 YR4 has a *cumulative impact probability* of 8.1e-6. Other ways of interpreting this number? A 0.00081% chance of impact. A significant difference from when it was estimated to have a 3.1% chance! This is why the stressing of variability is important.

[Velocity (km/s)]{style="font-size: 18px;font-weight: bold;"}: This is the calculated velocity of the asteroid relative to the Earth when it is at an effectively infinite distance, or not currently experiencing gravitational influence from the object, as represented in kilometers per second. This is also known as the *V*~∞~.

[Est. Diameter (km)]{style="font-size: 18px;font-weight: bold;"}: This is the best guess of the object's diameter in kilometers. We say "best guess" because asteroids are only observed when they reflect light back to us on Earth, and *visible light* can present as the same level of brightness from objects of vastly different sizes. Infrared light is also observed of NEOs to more accurately determine their possible diameter. As well, NEOs do not present as spheres like planets, but in various shapes, which further increases difficulty in accurately determining their diameter given the different "views" we might get of the object at its imaging moment.

[Palermo Scale (cum.)]{style="font-size: 18px;font-weight: bold;"}: This is the cumulative hazard rating according to the Palermo technical impact hazard scale. Generally, if a scale value is *less than* -2, there are likely minimal risks posed by the object. Scale values *between* 0 and -2 are worth careful consideration. Scale values that are *greater than* 0 warrant serious attention (and fortunately, nothing at current falls anywhere near this). The Palermo Scale is a logarithmic scale that is also designed to compare the likelihood of the object's detected possible impact with the average risk posed by objects of similar size (or larger) over the year range until the date of possible impact, which is known as the *background risk*. So, a Palermo Scale value of -2 means that object's possible hazard is 1% as likely as *any* other background risk of comparable size in the intervening years, while a value of 0 means the object is equally as much of a risk as other background risks, and a value of 2 indicates an event that is 10,000% more likely to occur than other background risks.

[Palermo Scale (max.)]{style="font-size: 18px;font-weight: bold;"}: This is the maximum possible measured Palermo Scale rating for a given object. This usually only presents when an object has more than one possible impact in its year range.

[Torino Scale (max.)]{style="font-size: 18px;font-weight: bold;"}: This is the maximum detected hazard rating according to the Torino impact hazard scale, based on the calculated impact probability and impact energy. The Torino Scale is defined only for potential impacts less than 100 years in the future. Unlike the Palermo Scale, the Torino Scale was developed to be more approachable to the public and is instead based on a range of 0-10, where: 0 indicates *no hazard* and is generally assigned to objects that will burn up in the atmosphere; 1 indicates *normal* behavior of the object with no discernible level of danger; 2-4 indicates the object *merits attention* by astronomers due to possible impact risk and subsequent damage; 5-7 indicates a *threatening* status from the object where it poses a potentially unprecedented threat; and 8-10, which indicate *certainty* of collision.

[Last Observed]{style="font-size: 18px;font-weight: bold;"}: This is the last time the object was observed and its data updated in Sentry.

# Potential Impacts Timeline

::: sidebar
## About This Graph

This graph shows a timeline of possible asteroid impacts based on current data from Sentry.

Each bubble represents a 10-year time window, beginning with the current decade (2020-2029). Bubble sizes increase based on overall number of possible asteroids with impact risks in those timeframes. Bubble sizes are not proportional.

**A note**: No data is provided after the year 2126 as Sentry only provides tracking information for objects with possible impacts in a 100-year window. Anomalies that do appear are excluded from this visualization but remain in the original dataset potentially as the result of prominent media presence, as is the case with the asteroid 101955 Bennu. The objects are as follows:

```{r}
#| label: after-2126-table
#| echo: false
#| results: asis

if (nrow(objects_after_2126_tbl) == 0) {
  cat("<div style='font-size:0.9em;opacity:0.8;'>None.</div>")
} else {
  knitr::kable(
    objects_after_2126_tbl,
    format = "html",
    col.names = c("Object", "Impact range"),
    escape = TRUE
  )
}
```

`r nrow(objects_after_2126_tbl)`

## Data Recency

```{r, echo=FALSE, results='asis'}
render_utc <- format(Sys.time(), tz = "UTC", "%Y-%m-%dT%H:%M:%SZ")

cat(paste0(
  "<span id='last-updated' data-utc='", render_utc, "'></span><br>",
  "<span id='next-refresh'></span>"
))
```

```{=html}
<script>
  (function () {
    // Read the UTC timestamp embedded by R
    const el = document.getElementById('last-updated');
    if (!el) return;
    
    const utc = el.getAttribute('data-utc');
    const last = new Date(utc);
    
    // Compute next refresh (daily at 14:00 UTC)
    const nextUtc = new Date();
    nextUtc.setUTCHours(14, 0, 0, 0);
    if (new Date() >= nextUtc) {
      nextUtc.setUTCDate(nextUtc.getUTCDate() + 1);
    }
    
    // Format in the viewer's local timezone
    const fmt = new Intl.DateTimeFormat(undefined, {
    dateStyle: 'long',
    timeStyle: 'short'
    });
    
    el.textContent = "This data was last updated on " + fmt.format(last) + ".";
    document.getElementById('next-refresh').textContent = 
      "Next scheduled update: " + fmt.format(nextUtc) + ".";
  })();
</script>
```
:::

```{r}
#| label: impacts-timeline-bubble-plot
#| warning: false

bin_width <- 10

impacts_parsed <- asteroid_table_typed |> 
  mutate(
    range = str_trim(range),
    range_ok = str_detect(range, "^\\d{4}\\s*-\\s*\\d{4}$")
  ) |> 
  filter(range_ok) |> 
  separate(
    range,
    into = c("start_year", "end_year"),
    sep = "\\s*-\\s*",
    convert = TRUE,
    remove = FALSE
  ) |> 
  filter(!is.na(start_year), !is.na(end_year), end_year >= start_year)

bin_start <- floor(min(impacts_parsed$start_year) / bin_width) * bin_width
bin_end <- ceiling(max(impacts_parsed$end_year) / bin_width) * bin_width
  
impacts_timeline <- tibble(
  bin_lo = seq(bin_start, bin_end - bin_width, by = bin_width),
  bin_hi = bin_lo + bin_width - 1
) |> 
  rowwise() |> 
  mutate(
    n_objects = sum(impacts_parsed$start_year <= bin_hi & impacts_parsed$end_year >= bin_lo),
    bin_mid = (bin_lo + bin_hi) / 2,
    bin_label = paste0(bin_lo, "-", bin_hi)
  ) |> 
  ungroup() |> 
  filter(n_objects > 0) |> 
  arrange(bin_mid) |> 
  mutate(
    y = 0,
    side = if_else(row_number() %% 2 == 0, "below", "above"),
    y_label = if_else(side == "above", 0.30, -0.30),
    y_line_end = if_else(side == "above", 0.27, -0.27),
    label = paste0(bin_label, "\n(n = ", n_objects, ")")
  )

xmin <- min(impacts_timeline$bin_lo)
xmax <- max(impacts_timeline$bin_hi)

ggplot(impacts_timeline, aes(x = bin_mid, y = y)) +
  geom_segment(
    aes(x = xmin, xend = xmax, y = 0, yend = 0),
    inherit.aes = FALSE,
    linewidth = 1.1,
    arrow = arrow(length = unit(0.22, "cm"), type = "closed")
  ) +
  geom_segment(
    aes(xend = bin_mid, yend = y_line_end),
    linewidth = 0.35, alpha = 0.35
  ) +
  geom_point(
    aes(size = n_objects),
    shape = 21,
    fill = "#56B4E9",
    color = "#56B4E9",
    alpha = 0.65
  ) +
  geom_text(
    aes(y = y_label, label = label),
    size = 2.5,
    lineheight = 0.95,
    angle = 45,
    hjust = 0.5,
    vjust = 0.5) +
  scale_x_continuous(
    name = NULL,
    breaks = NULL,
    labels = NULL,
    expand = expansion(mult = c(0.05, 0.10))
  ) +
  scale_y_continuous(
    name = NULL,
    breaks = NULL
  ) +
  scale_size(range = c(4, 20), guide = "none") +
  coord_cartesian(clip = "off") +
  theme_void(base_size = 14) +
  theme(
    plot.margin = margin(25, 25, 25, 25),
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20))
    ) +
  labs(
    title = "Timeline of Possible NEO Impacts Between 2026 and 2289"
  )
```

# Palermo Scale Risk Distribution

```{r}
#| label: palermo-bar-chart
#| warning: false
#| fig-cap: "Distribution of monitored asteroids across Palermo Scale risk categories"

asteroid_risk <- asteroid_table_typed |>
  mutate(
    ps_cum_cat = case_when(
      ps_cum > 0 ~ ">0",
      ps_cum <= 0 & ps_cum > -2 ~ ">0 to =-2",
      ps_cum <= -2 & ps_cum > -4 ~ ">-2 to =-4",
      ps_cum <= -4 & ps_cum > -6 ~ ">-4 to =-6",
      ps_cum <= -6 ~ "<=-6",
      TRUE ~ NA_character_
    ),
    ps_max_cat = case_when(
      ps_max > 0 ~ ">0",
      ps_max <= 0 & ps_max > -2 ~ ">0 to =-2",
      ps_max <= -2 & ps_max > -4 ~ ">-2 to =-4",
      ps_max <= -4 & ps_max > -6 ~ ">-4 to =-6",
      ps_max <= -6 ~ "<=-6",
      TRUE ~ NA_character_
    )
  )

risk_long <- asteroid_risk |>
  select(ps_cum_cat, ps_max_cat) |>
  pivot_longer(
    cols = everything(),
    names_to = "scale",
    values_to = "category"
  ) |>
  filter(!is.na(category))

risk_long$category <- factor(
  risk_long$category,
  levels = c(">0", ">0 to =-2", ">-2 to =-4", ">-4 to =-6", "<=-6")
)

risk_counts <- risk_long |>
  count(scale, category)

ggplot(risk_counts,
       aes(x = category, y = n, fill = scale)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = comma) +
  labs(
    x = "Palermo Scale Risk Category",
    y = "Number of Asteroids",
    fill = "Palermo Scale",
    title = "Relative Risk Distribution of Monitored Near-Earth Asteroids",
    subtitle = "Comparison of cumulative (ps_cum) and maximum (ps_max) Palermo Scale values"
  ) +
  theme_minimal(base_size = 12)

```

# Sentry Objects Table

```{r}
#| label: sentry-object-table
#| echo: false

obj_tbl <- asteroid_table_typed |> 
  transmute(
    designation,
    fullname,
    range,
    n_imp,
    ip,
    v_inf,
    diameter,
    ps_cum,
    ps_max,
    ts_max,
    last_obs
  )

datatable(
  obj_tbl,
  rownames = FALSE,
  selection = "single",
  colnames = c(
    "Designation",
    "Object",
    "Impact Year Range",
    "Potential Impacts",
    "Impact Probability (cumulative)",
    "Velocity (km/s)",
    "Est. Diameter (km)",
    "Palermo Scale (cum.)",
    "Palermo Scale (max.)",
    "Torino Scale (max.)",
    "Last Observed"
  ),
  options = list(
    pageLength = 10,
    paging = TRUE,
    info = TRUE,
    searching = TRUE,
    lengthMenu = list(c(10,25,50,100), c("10", "25", "50", "100")),
    lengthChange = TRUE,
    autoWidth = TRUE,
    scrollX = TRUE,
    dom = "<'dt-top'lfr>t<'dt-bottom'ip>",
    pagingType = "simple_numbers",
    columnDefs = list(list(targets = 0, visible = FALSE))
  ),
  callback = DT::JS("
    table.on('click.dt', 'tr', function() {
      var row = table.row(this);
      if (!row || !row.data()) return;
      
      var data = row.data();
      var des = data[0];
      
      var panel = document.getElementById('sentry-details');
      panel.innerHTML = '<div style=\"opacity:.8;\">Loading details for <b>' + des + '</b>...</div>';
      
      var url = 'https://ssd-api.jpl.nasa.gov/sentry.api?des=' + encodeURIComponent(des);
      
      fetch(url)
        .then(r => r.json())
        .then(payload => {
          if (payload.error) {
            panel.innerHTML = '<div style=\"color:#b00020;\">API error: ' + payload.error + '</div>';
            return;
          }
          
          var s = (payload.summary && payload.summary.length) ? payload.summary[0] : null;
          if (!s) {
            panel.innerHTML = '<div>No summary available for ' + des + '.</div>';
            return;
          }
          
          var html = '';
          html += '<h3 style=\"margin-top:0;\">' + (s.fullname || des) + '</h3>';
          html += '<div style=\"display:grid;grid-template-columns:1fr 1fr;gap:8px;\">';
          html += '<div><b>Year range:</b> ' + (s.range ?? '') + '</div>';
          html += '<div><b>Potential impacts:</b> ' + (s.n_imp ?? '') + '</div>';
          html += '<div><b>Cumulative IP:</b> ' + (s.ip ?? '') + '</div>';
          html += '<div><b>V∞ (km/s):</b> ' + (s.v_inf ?? '') + '</div>';
          html += '<div><b>Diameter (km):</b> ' + (s.diameter ?? '') + '</div>';
          html += '<div><b>Torino max:</b> ' + (s.ts_max ?? '') + '</div>';
          html += '<div><b>Palermo cum:</b> ' + (s.ps_cum ?? '') + '</div>';
          html += '<div><b>Palermo max:</b> ' + (s.ps_max ?? '') + '</div>';
          html += '</div>';
          
          var impacts = payload.data || [];
          html += '<h4 style=\"margin-top:14px;\">Impact events (sample)</h4>';
          
          if (!impacts.length) {
            html += '<div style=\"opacity:.8;\">No impact-event rows returned.</div>';
          } else {
            var head = Object.keys(impacts[0]);
            var showCols = ['date', 'ip', 'ps', 'ts', 'sigma', 'energy']; // best-effort
            var cols = head.filter(h => showCols.includes(h)).slice(0, 6);
            if (!cols.length) cols = head.slice(0, 6);

            html += '<table style=\"width:100%;border-collapse:collapse;\">';
            html += '<thead><tr>' + cols.map(c => '<th style=\"text-align:left;border-bottom:1px solid #ddd;padding:4px;\">' + c + '</th>').join('') + '</tr></thead>';
            html += '<tbody>';

            impacts.slice(0, 10).forEach(row => {
              html += '<tr>' + cols.map(c => '<td style=\"border-bottom:1px solid #f0f0f0;padding:4px;\">' + (row[c] ?? '') + '</td>').join('') + '</tr>';
            });

            html += '</tbody></table>';
            html += '<div style=\"margin-top:6px;opacity:.75;font-size:.9em;\">Showing 10 rows; API returned ' + impacts.length + '.</div>';
          }
          
          panel.innerHTML = html;
        })
        .catch(err => {
          panel.innerHTML = '<div style=\"color:#b00020;\">Fetch failed: ' + err + '</div>';
        });
      });
  ")
)
```
